---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/",
  out.width = "100%"
)
```

# elbow

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/ahasverus/elbow.svg?branch=master)](https://travis-ci.com/ahasverus/elbow)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ahasverus/elbow?branch=master&svg=true)](https://ci.appveyor.com/project/ahasverus/elbow)
[![CRAN\_Status\_Badge](https://www.r-pkg.org/badges/version/elbow)](https://cran.r-project.org/package=elbow)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](code_of_conduct.md)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Project Status: Stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
<!-- badges: end -->

The goal of the package `elbow` is to implement the Elbow (or knee of a curve)
method to detect the inflection point of a concave curve. More information on
this method can be found on [Wikipedia](https://en.wikipedia.org/wiki/Elbow_method_(clustering)).


## Installation

You can install the `elbow` package from [GitHub](https://github.com/ahasverus/elbow) with:

```{r echo = TRUE, eval = FALSE}
devtools::install_github("ahasverus/elbow", build_vignettes = TRUE)
library(elbow)
```


## Getting started

Visit the [**vignette**](https://nicolascasajus.fr/elbow/articles/introduction.html)
for a complete example. The vignette is also available by using:

```{r echo = TRUE, eval = FALSE}
browseVignettes("elbow")
```


## Code of Conduct

Please note that the `elbow` project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
