---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%"
)
```

# elbow

<!-- badges: start -->
<!-- badges: end -->

The goal of the package `elbow` is to implement the Elbow (or knee of a curve)
method to detect the inflection point of a concave curve. More information on
this method can be found on [Wikipedia](https://en.wikipedia.org/wiki/Elbow_method_(clustering)).

## Installation

You can install the dev version of `elbow` from [GitHub](https://github.com/ahasverus/elbow) with:

```{r echo=TRUE, eval=FALSE}
devtools::install_github("ahasverus/elbow")
```

## Getting started

Browse the example of the function `elbow::elbow()` for a complete usage:

```{r echo=TRUE, eval=FALSE}
library(elbow)
?elbow
```

## Code of Conduct

Please note that the `elbow` project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
