---
title: "The `{elbow}` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "## ",
  fig.align  = "center",
  fig.height = 6,
  fig.width  = 6,
  out.height = "100%",
  out.width  = "100%",
  dpi        = 300
)
```

```{r params, echo = FALSE, eval = TRUE}
library(elbow)

param <- list(
  mar      = c(2.5, 2.5, 1.5, 1.5),
  family   = "serif",
  cex.axis = 0.85,
  mgp      = c(2, .15, 0),
  tcl      = -0.25,
  fg       = "#666666",
  col      = "#666666",
  col.axis = "#666666"
)
```


The goal of the package **`{elbow}`** is to implement the Elbow (or knee of a curve)
method to detect the inflection point of a concave curve. More information on
this method can be found below and on [Wikipedia](https://en.wikipedia.org/wiki/Elbow_method_(clustering)).

<br />

## Getting started

We can install the package **`{elbow}`** from [GitHub](https://github.com/ahasverus/elbow) with:

```{r install, echo = TRUE, eval = FALSE}
devtools::install_github("ahasverus/elbow")

library(elbow)
```

<br />

Let's take a look at the package content:

```{r load, echo = TRUE, eval = TRUE}
ls("package:elbow")
```

The package contains two elements:

- **`elbow()`** - the only function of the package used to detect inflection point;
- **`profiles`** - a reproductible example.

<br />

## Reproductible example

First, we will load the **`profiles`** dataset and print a summary of its content:

```{r usage, echo = TRUE, eval = TRUE}
data(profiles)

str(profiles)

head(profiles)
```

<br />

The **`profiles`** dataset is a data frame with 31 rows and the five following variables:

| Variable name              | Description                                           |
|:---------------------------|:------------------------------------------------------|
| **`x`**                    | A sequence from 0 to 30 (x-Axis)                      |
| **`concave_down_pos_slo`** | 1^st^ profile - Concave down with positive slope      |
| **`concave_down_neg_slo`** | 2^nd^ profile - Concave down with negative slope      |
| **`concave_up_pos_slo`**   | 3^rd^ profile - Concave up with positive slope        |
| **`concave_up_neg_slo`**   | 4^th^ profile - Concave up with negative slope        |

Let's plot these four concave curves along the **`x`** sequence:

<br />

```{r plot1, echo = FALSE, eval = TRUE}
par(param)
par(mfrow = c(2, 2))

for (i in 2:5) {

  xy <- profiles[ , c(1, i)]
  plot(xy, xlim = c(0, 30), ylim = c(0, 1), ann = FALSE, axes = FALSE, type = "n")
  grid()
  box()

  par(mgp = c(2, 0.00, 0))
  axis(1, lwd = 0)

  par(mgp = c(2, 0.25, 0))
  axis(2, lwd = 0, las = 1)

  mtext(side = 1, cex = 1, line = 1.25, text = expression("x"))
  mtext(side = 2, cex = 1, line = 1.45, text = expression("f(x)"))

  points(xy, type = "b", pch = 19, col = "black")

  rect(7, .41, 23, .59, col = "white", border = NA)

  label <- gsub("_", " ", colnames(xy)[2])
  label <- gsub("concave", "Concave", label)
  label <- gsub("down", "Down", label)
  label <- gsub("up", "Up", label)
  label <- gsub("pos slo", "\n(positive slope)", label)
  label <- gsub("neg slo", "\n(negative slope)", label)

  text(15, 0.5, label, font = 2)

  if (i %in% c(2, 4)) {

    text(-1, 0.95, LETTERS[i-1], font = 2, pos = 4)

  } else {

    text(31, 0.95, LETTERS[i-1], font = 2, pos = 2)
  }
}
```

<br />

**NB.** In Clustering Analysis or Principal Component Analysis, we frequently meet
the profiles **A** (a quantity increasing as the number of groups increases) and
**D** (a quantity decreasing as the number of groups increases).

<br />

## The Elbow algorithm

We are going to detect the inflection point of the profile **A** (Concave down
with a positive slope). In Clustering Analysis, the x-Axis may represent the number
of groups and the y-Axis the explained variance (R^2^).

The idea behind the Elbow method is to maximize a quantity (benefits) while reducing
the costs (number of groups). Consequently, the inflection point will be the point
from which the benefits become lower than the costs.

**NB.** In the profile **D** (Concave up with a negative slope) the objective is to
minimize the quantity while reducing the costs.

<br />

How does it work?

```{r plot2, echo = FALSE, eval = TRUE}
par(param)
par(mfrow = c(2, 2))

xy <- profiles[ , c(1, 2)]
ipoint <- elbow(data = xy, plot = FALSE)

plot(xy, xlim = c(0, 30), ylim = c(0, 1), ann = FALSE, axes = FALSE, type = "n")
grid()
box()

par(mgp = c(2, 0.00, 0))
axis(1, lwd = 0)

par(mgp = c(2, 0.25, 0))
axis(2, lwd = 0, las = 1)

mtext(side = 1, cex = 1, line = 1.25, text = expression("x"))
mtext(side = 2, cex = 1, line = 1.45, text = expression("f(x)"))

points(xy, type = "b", pch = 19, col = "black")
text(-1, 0.95, "(1)", font = 2, pos = 4)


plot(xy, xlim = c(0, 30), ylim = c(0, 1), ann = FALSE, axes = FALSE, type = "n")
grid()
box()

par(mgp = c(2, 0.00, 0))
axis(1, lwd = 0)

par(mgp = c(2, 0.25, 0))
axis(2, lwd = 0, las = 1)

mtext(side = 1, cex = 1, line = 1.25, text = expression("x"))
mtext(side = 2, cex = 1, line = 1.45, text = expression("f(x)"))

points(xy, type = "b", pch = 19, col = "#aaaaaa66")

for (i in 1:nrow(xy)) {
  rect(
    xleft   = ipoint$data[i, "x"] - .5,
    ybottom = 0,
    xright  = ipoint$data[i, "x"] + .5,
    ytop    = ipoint$data[i, "constant"],
    border  = "white",
    col     = "black"
  )
}

text(-1, 0.95, "(2)", font = 2, pos = 4)



plot(xy, xlim = c(0, 30), ylim = c(0, 1), ann = FALSE, axes = FALSE, type = "n")
grid()
box()

par(mgp = c(2, 0.00, 0))
axis(1, lwd = 0)

par(mgp = c(2, 0.25, 0))
axis(2, lwd = 0, las = 1)

mtext(side = 1, cex = 1, line = 1.25, text = expression("x"))
mtext(side = 2, cex = 1, line = 1.45, text = expression("f(x)"))

points(xy, type = "b", pch = 19, col = "#aaaaaa66")

for (i in 1:nrow(xy)) {
  rect(
    xleft   = ipoint$data[i, "x"] - .5,
    ybottom = 0,
    xright  = ipoint$data[i, "x"] + .5,
    ytop    = ipoint$data[i, "constant"],
    border  = "white",
    col     = "#aaaaaa66"
  )

  if (!(i %in% c(1, nrow(xy)))) {
    arrows(
      x0 = ipoint$data[i, "x"],
      y0 = ipoint$data[i, "constant"],
      x1 = ipoint$data[i, "x"],
      y1 = ipoint$data[i, 2],
      length = 0.025,
      angle  = 20,
      code   = 3,
      col = "black"
    )
  }
}

text(-1, 0.95, "(3)", font = 2, pos = 4)


plot(xy, xlim = c(0, 30), ylim = c(0, 1), ann = FALSE, axes = FALSE, type = "n")
grid()
box()

par(mgp = c(2, 0.00, 0))
axis(1, lwd = 0)

par(mgp = c(2, 0.25, 0))
axis(2, lwd = 0, las = 1)

mtext(side = 1, cex = 1, line = 1.25, text = expression("x"))
mtext(side = 2, cex = 1, line = 1.45, text = expression("f(x)"))

points(xy, type = "b", pch = 19, col = "#aaaaaa66")

for (i in 1:nrow(xy)) {
  rect(
    xleft   = ipoint$data[i, "x"] - .5,
    ybottom = 0,
    xright  = ipoint$data[i, "x"] + .5,
    ytop    = ipoint$data[i, "constant"],
    border  = "white",
    col     = "#aaaaaa66"
  )
}

polygon(
  x      = c(ipoint$data[ , 1], ipoint$data[1, 1]),
  y      = c(ipoint$data[ , "benefits"], ipoint$data[1, "benefits"]),
  col    = "black",
  border = "black"
)

opti <- ipoint$dat[ipoint$dat[ , 1] == ipoint[[1]], c(1, 4)]
points(opti, type = "b", pch = 19, col = "#890000", cex = 1.5)


text(-1, 0.95, "(4)", font = 2, pos = 4)

```

From the profile **(1)**, we apply a constant increase along the
x-Axis to reach the maximum value on the y-Axis **(2)**. Then, for each value on
the x-Axis, we compute the difference between the two series **(3)** to generate the
profile **(4)** (net benefits). The inflection point, in this case, corresponds
to the maximum value on this new profile (red dot).


<br />

## Usage of **`elbow()`**

This algorithm is implemented in the function **`elbow::elbow()`** which takes two
arguments:

- **`data`** - a two-columns data frame (x and y respectively);
- **`plot`** - a boolean. If `TRUE` (default) the curves are plotted.

<br />

Let's apply this function to detect the inflection point of the profile **A**.

```{r plot3, echo = TRUE, eval = TRUE}
dopo <- profiles[ , c("x", "concave_down_pos_slo")]

ipoint <- elbow(data = dopo)
```

<br />

What is the returned object?

```{r content, echo = TRUE, eval = TRUE}
class(ipoint)

names(ipoint)

print(ipoint)
```

The element `ipoint$data` is returned to reproduce the graphic. The one we are
interested in is **`ipoint$x_selected`** which returns the coordinate on the x-Axis
(not the position on the data frame) corresponding to the inflection point.

```{r axis, echo = TRUE, eval = TRUE}
ipoint$"x_selected"

ipoint[["data"]][ipoint[["data"]][ , "x"] == ipoint[["x_selected"]], 1:2]
```

<br />

What about the other profiles?

```{r plot4, echo = TRUE, eval = TRUE}
profs <- colnames(profiles)[-1]

par(mfrow = c(2, 2))

for (prof in profs) {
  elbow(profiles[ , c("x", prof)])
}
```

<br />

## Code of Conduct

Please note that the `elbow` project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

<br />

Last updated: `r format(Sys.time(), "%Y/%m/%d")`
